---
title: "FIN511 Week 1"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

First code chunk used to load the necessary libraries in R
if you don't have these libraries, please uncomment the code and install them or install them in R-Studio.

```{r}
# load libraries

# install readxl and tidyverse (or just ggplot2 and dyplr)
# install.packages("readxl","ggplot2","dplyr")

library(ggplot2)
library(readxl)
library(dplyr)
library(reshape2)
```

The following code chunk shows a very "dumb" way to load excel spreadsheets in R directly, skipping lines
if necessary, like in this case and use the function "head" to see if the spreadsheet is loaded properly


```{r}
# Load all the datasets and separate sheets, clean up the data and leave a single row for column names.
# You will still need the original excel sheet to read the instructions at the top or additional informations
# modify the files paths to match your needs
# I didn't set working directories or anything else to keep the customization to a minimum

df1_mom_returns <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/213722-XLS-ENG.xlsx", sheet = 2, skip = 2)

df1_portfolio_strategies <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/213722-XLS-ENG.xlsx", sheet = 3, skip = 3)

rt_vol_risky_and_rf <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/Return-Volatility-EXAMPLES.xlsx", sheet = 1, skip = 1)

rt_vol_risky_assets <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/Return-Volatility-EXAMPLES.xlsx", sheet = 2, skip = 1)

head(df1_mom_returns)
```
Lets replicate the "two risky assets" spreadsheet

Step 1, let's get all the information together and organize it in a dataframe (a structure similar to a spreadsheet)
```{r}
# Risk Free rate
RF <- 3
# The large Weight is a sequence that goes from -1 to 2 in 0.01 steps
Large_Weight <- seq(-1,2,0.01)
# The Small Weight is simply 1 - Large Weight
Small_Weight <- 1 - Large_Weight

# Large return
LR <- 8
# Small return
SR <- 15

# Correlation (using CR because cor is an R funcion therefore reserved word)
CR <- 0.4

# Large Standard Deviation
LSD <- 25
# Small Standard Deviation 
SSD <- 50

# lets create the dataframe. Note, R automatically recycle variables to fit the lenght of the dataframe
# therefore is important that the first variable is the correct lenght.

df <- data.frame("RF" = RF ,"Large_Weight" = Large_Weight, "Small_Weight" = Small_Weight, "LR" = LR, "SR" = SR,
                 "CR" = CR, "LSD" = LSD, "SSD" = SSD)

# We could have also avoided all the variable declarations and just do:
# df <- data.frame("Large_Weight" = seq(-1,2,0.01), "Small_Weight" = 1 - seq(-1,2,0.01), "LR" = 8)

# Now let's add the calculated columns (note that column can be added just using df "$" name_of_the_column)
# and we use the dollar sign to select the column by name as well
# Important Don't use spaces in the name of columns or variables or anything in general, it's just bad 
# practice at best and in most cases the software will give you an error
df$Portfolio_SD <- sqrt(df$Large_Weight^2 * df$LSD^2 + df$Small_Weight^2 * df$SSD^2 + 
                          2*df$CR*df$Large_Weight*df$LSD*df$Small_Weight*df$SSD)

df$Portfolio_Return <- df$Large_Weight*df$LR + df$Small_Weight*df$SR

df$Sharpe_Ratio <- (df$Portfolio_Return - df$RF)/df$Portfolio_SD

# let's verify that our dataframe is created correctly
head(df)
tail(df)
```

Now that the data has been created, let's create the charts in ggplot2
protip: download the ggplot2 cheat sheet https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf

```{r}
# Let's use GGPLOT to create our charts, it gives us more customization abilities
# for example we can encode the sharpe ratio as color along the curve

ggplot(df, aes(x=Portfolio_SD, y=Portfolio_Return)) +
  geom_point(aes(color=Sharpe_Ratio)) +
  xlab("Portfolio Standard Deviation") +
  ylab("Portfolio rate or return (in %)")
  
```

Let's create the second chart without the option of shorting
```{r}
# Careful, I'm using the filter function from the package dplyr, which is different from the standard
# one, therefore, load dyplr as library.

ggplot(filter(df, Large_Weight < 1, Small_Weight < 1), aes(x=Portfolio_SD, y=Portfolio_Return)) +
  geom_point(aes(color=Sharpe_Ratio)) +
  xlab("Portfolio Standard Deviation") +
  ylab("Portfolio rate or return (in %)")
```

How to generate the answers automatically

First let's create a second dataframe with the modified parameters and call it df2
```{r}
# Risk Free rate
RF <- 3
# The large Weight is a sequence that goes from -1 to 2 in 0.01 steps
Large_Weight <- seq(-1,2,0.01)
# The Small Weight is simply 1 - Large Weight
Small_Weight <- 1 - Large_Weight

# Large return
LR <- 8
# Small return
SR <- 15

# Correlation (using CR because cor is an R funcion therefore reserved word)
CR <- -0.8

# Large Standard Deviation
LSD <- 25
# Small Standard Deviation 
SSD <- 50

# lets create the dataframe. Note, R automatically recycle variables to fit the lenght of the dataframe
# therefore is important that the first variable is the correct lenght.

df2 <- data.frame("RF" = RF ,"Large_Weight" = Large_Weight, "Small_Weight" = Small_Weight, "LR" = LR, "SR" = SR,
                 "CR" = CR, "LSD" = LSD, "SSD" = SSD)

# We could have also avoided all the variable declarations and just do:
# df <- data.frame("Large_Weight" = seq(-1,2,0.01), "Small_Weight" = 1 - seq(-1,2,0.01), "LR" = 8)

# Now let's add the calculated columns (note that column can be added just using df "$" name_of_the_column)
# and we use the dollar sign to select the column by name as well
# Important Don't use spaces in the name of columns or variables or anything in general, it's just bad 
# practice at best and in most cases the software will give you an error
df2$Portfolio_SD <- sqrt(df2$Large_Weight^2 * df2$LSD^2 + df2$Small_Weight^2 * df2$SSD^2 + 
                          2*df2$CR*df2$Large_Weight*df2$LSD*df2$Small_Weight*df2$SSD)

df2$Portfolio_Return <- df2$Large_Weight*df2$LR + df2$Small_Weight*df2$SR

df2$Sharpe_Ratio <- (df2$Portfolio_Return - df2$RF)/df2$Portfolio_SD

# let's verify that our dataframe is created correctly
head(df2)
tail(df2)
```

Let see the modified chart

```{r}
# Let's use GGPLOT to create our charts, it gives us more customization abilities
# for example we can encode the sharpe ratio as color along the curve


ggplot(df2, aes(x=Portfolio_SD, y=Portfolio_Return)) +
  geom_point(aes(color=Sharpe_Ratio)) +
  xlab("Portfolio Standard Deviation") +
  ylab("Portfolio rate or return (in %)")
  
```
And now overlay them. In R this is not very automated, please follow the code
This example is simplified by the fact that the portfolio rate of return does not depend on 
the correlation coefficent but only by the weights, so we can use it "as is" for our Y axis

```{r}
# first we need to create a dataframe with the information we need. 
# the portfolio rate of return is our Y and cor_04 and cor_min08 are the standard deviations

# create a simple dataframe with only the variables of interest (this step is not necessary but it is easier to verify)
overlay_df <- data.frame(Y=df$Portfolio_Return, cor_04=df$Portfolio_SD, cor_min08=df2$Portfolio_SD)

# melt the dataframe to stack the variables in a single column
overlay_df.m <- melt(overlay_df, id.vars = "Y", measure.vars = c("cor_04", "cor_min08"))

# double check the stacking is correct 
head(overlay_df.m)
tail(overlay_df.m)


# plot reversing the axis (we put Y where X is expected and so on)
ggplot(overlay_df.m, aes(value, Y, colour = variable)) + 
  geom_point() + 
  xlab("Portfolio Standard Deviation") +
  ylab("Portfolio rate or return (in %)")
```

```{r}

```

Question 1: Find the composition of the portfolio at minimum variance 
```{r}
# the function filter from dyplr is very useful in this case. 
filter(df2, Portfolio_SD == min(Portfolio_SD))

```

Question 2: 
```{r}

```

Question 3 shows something that is very common... what if there isn't a portfolio return exactly equal to 18.
Note: -check the numbers, I'm giving the template, not the solution
can be approached in 2 ways. First, a brute force approach using the filter function from dyplr
Second, we can calculate exactly the 

Question 3:
```{r}
# Test if there is a return exactly equal to 18.
filter(df2, Portfolio_Return == 18)

# then we just filter the portfolio around the value and we select the correct row
filter(df2, Portfolio_Return > 17.5 & Portfolio_Return < 18.5)

# or we just use a neat function "which.min" which returns the row where the absolute value
# of the difference between the Portfolio_Return and 20 is at its minimum

row.number <- which.min(abs(df2$Portfolio_Return - 18))

# and we verify that is correct.
df2[row.number,]

```



```{r}
```

```{r}

```