---
title: "FIN511 Week 2"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

First code chunk used to load the necessary libraries in R
if you don't have these libraries, please uncomment the code and install them or install them in R-Studio.

```{r}
# load libraries

# install readxl and tidyverse (or just ggplot2 and dyplr)
# install.packages("readxl","ggplot2","dplyr")

library(ggplot2)
library(readxl)
library(reshape2)
```

The following code chunk shows a very "dumb" way to load excel spreadsheets in R directly, skipping lines
if necessary, like in this case and use the function "head" to see if the spreadsheet is loaded properly


```{r}
# Load all the datasets and separate sheets, clean up the data and leave a single row for column names.
# You will still need the original excel sheet to read the instructions at the top or additional informations
# modify the files paths to match your needs
# I didn't set working directories or anything else to keep the customization to a minimum

# This excel spreadsheet is not very clean, data starts on row 4, there is an empty column and on column J there are  calculations.
# Let see how to clean it up.


df.small_value <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/SmallValue_1927_2014.xlsx", sheet = 1, skip = 3)

# we observe that column 6 is empty and 9 and 10 are not useful
head(df.small_value)

# let's clean up: nothing easier, we just say that our dataframe is the same dataframe minus column 6,9 and 10
# just remember that the indexes in the square brakets are [rows, columns] and notice the "-c" that indicates 
# minus this list, otherwise we would have kept only column 6, 9, 10

df.small_value <- df.small_value[,-c(6,9,10)]

# double check
head(df.small_value)
```
## Evaluation of Small-Value Stock investment strategy

here we are going to compute the required values and present them.

```{r}
# computation of the values, rounded to the third digit
av.return.in.pct = round(mean(df.small_value$SmallValue),digits = 3)
av.exc.ret.in.pct = round(mean(df.small_value$SmallValue_RF),digits = 3)
us.stock.mkt.exc.ret = round(mean(df.small_value$Mkt_RF), digits = 3)
small.cap.minus.large.cap = round(mean(df.small_value$SMB), digits = 3)
value.min.growth = round(mean(df.small_value$HML), digits = 3)

```

Now we can report:

Small-Value Stocks Return(SmallValue): `r av.return.in.pct`%

Small-Value Stocks Excess Return (SmallValue_RF): `r av.exc.ret.in.pct`%

US Stock Market Excess Return (Mkt_RF): `r us.stock.mkt.exc.ret`%

Small-cap minus Large-cap Stocks (SMB): `r small.cap.minus.large.cap`%

Value minus Growth Stocks (HML): `r value.min.growth`%

## Small value stocks CAPM
```{r}
# step 1 let build the model (lm is the function for linear models, and the regression is a linear model)
model <- lm(SmallValue_RF ~ Mkt_RF ,data=df.small_value)

# step 2 let see the model and extract the coefficients
model

alpha <- model$coefficients[[1]]
beta <- model$coefficients[[2]]

# step 3 let verify r-squared
r_squared <- summary(model)$r.squared

# let show the anova table to verify if the model is valid 
anova(model)

# and the summary of the model
summary(model)
```

Average return (in %): `r av.return.in.pct`%

Average excess return (in %): `r av.exc.ret.in.pct`%

CAPM alpha (in %): `r alpha`

CAPM market beta: `r beta`

R-squared: `r r_squared` 


## 3-Factor model
```{r}
# step 1 let build the model
model.3 <- lm(SmallValue_RF ~ Mkt_RF + SMB + HML ,data=df.small_value)

# step 2 let see the model and extract the coefficients
model.3

# notice the index values reflect the order in which the variables are put in the equation
# except for alpha (intercept) which will be 1
alpha.3 <- model.3$coefficients[[1]]
beta.3.market <- model.3$coefficients[[2]]
beta.3.size <- model.3$coefficients[[3]]
beta.3.value <- model.3$coefficients[[4]]
# step 3 let verify r-squared
r_squared.3 <- summary(model.3)$r.squared

# let show the anova table to verify if the model is valid 
anova(model.3)

# and the summary of the model
summary(model.3)
```

3-Factor alpha (in %): `r alpha.3`

3-Factor market beta: `r beta.3.market`

3-Factor size beta: `r beta.3.size`

3-Factor value: `r beta.3.value`

R-squared: `r r_squared.3`


### Load the Three mystery securities

```{r}
df.3.myst <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/ThreeMysterySecurities.xlsx", sheet = 1, skip = 3)

df.3.myst <- df.3.myst[,-6]

head(df.3.myst)

```

### generate basic stats and CAPM regression for mystery security 1
```{r}
# computation of the values, rounded to the third digit
mys.1.av.ret = round(mean(df.3.myst$`Security 1`),digits = 3)
mys.1.av.ex.ret = round(mean(df.3.myst$`Security 1` - df.3.myst$RF),digits = 3)

# step 1 let build the model (lm is the function for linear models, and the regression is a linear model)
model.m.1 <- lm(`Security 1`~ Mkt_RF ,data=df.3.myst)

# step 2 let see the model and extract the coefficients
model.m.1

mys.1.CAPM.alpha <- model.m.1$coefficients[[1]]
mys.1.CAPM.beta <- model.m.1$coefficients[[2]]

# step 3 let verify r-squared
mys.1.r.squared <- summary(model.m.1)$r.squared

# let show the anova table to verify if the model is valid 
anova(model.m.1)

# and the summary of the model
summary(model.m.1)


```

