---
title: "FIN511 Week 1"
output:
  html_document:
    df_print: paged
---

First code chunk used to load the necessary libraries in R
if you don't have these libraries, please uncomment the code and install them or install them in R-Studio.

```{r}
# load libraries

# install readxl and tidyverse (or just ggplot2 and dyplr)
# install.packages("readxl","ggplot2","dplyr")

library(ggplot2)
library(readxl)
library(dplyr)
```

The following code chunk shows a very "dumb" way to load excel spreadsheets in R directly, skipping lines
if necessary, like in this case and use the function "head" to see if the spreadsheet is loaded properly


```{r}
# Load all the datasets and separate sheets, clean up the data and leave a single row for column names.
# You will still need the original excel sheet to read the instructions at the top or additional informations
# modify the files paths to match your needs
# I didn't set working directories or anything else to keep the customization to a minimum

df1_mom_returns <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/213722-XLS-ENG.xlsx", sheet = 2, skip = 2)

df1_portfolio_strategies <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/213722-XLS-ENG.xlsx", sheet = 3, skip = 3)

rt_vol_risky_and_rf <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/Return-Volatility-EXAMPLES.xlsx", sheet = 1, skip = 1)

rt_vol_risky_assets <- read_xlsx("/Users/ataru074/Desktop/Education/MBA/03 2020 Fall/FIN511 Investments/Return-Volatility-EXAMPLES.xlsx", sheet = 2, skip = 1)

head(df1_mom_returns)
```
Lets replicate the "two risky assets" spreadsheet

Step 1, let's get all the information together and organize it in a dataframe (a structure similar to a spreadsheet)
```{r}
# Risk Free rate
RF <- 3
# The large Weight is a sequence that goes from -1 to 2 in 0.01 steps
Large_Weight <- seq(-1,2,0.01)
# The Small Weight is simply 1 - Large Weight
Small_Weight <- 1 - Large_Weight

# Large return
LR <- 8
# Small return
SR <- 15

# Correlation (using CR because cor is an R funcion therefore reserved word)
CR <- 0.4

# Large Standard Deviation
LSD <- 25
# Small Standard Deviation 
SSD <- 50

# lets create the dataframe. Note, R automatically recycle variables to fit the lenght of the dataframe
# therefore is important that the first variable is the correct lenght.

df <- data.frame("RF" = RF ,"Large_Weight" = Large_Weight, "Small_Weight" = Small_Weight, "LR" = LR, "SR" = SR,
                 "CR" = CR, "LSD" = LSD, "SSD" = SSD)

# We could have also avoided all the variable declarations and just do:
# df <- data.frame("Large_Weight" = seq(-1,2,0.01), "Small_Weight" = 1 - seq(-1,2,0.01), "LR" = 8)

# Now let's add the calculated columns (note that column can be added just using df "$" name_of_the_column)
# and we use the dollar sign to select the column by name as well
# Important Don't use spaces in the name of columns or variables or anything in general, it's just bad 
# practice at best and in most cases the software will give you an error
df$Portfolio_SD <- sqrt(df$Large_Weight^2 * df$LSD^2 + df$Small_Weight^2 * df$SSD^2 + 
                          2*df$CR*df$Large_Weight*df$LSD*df$Small_Weight*df$SSD)

df$Portfolio_Return <- df$Large_Weight*df$LR + df$Small_Weight*df$SR

df$Sharpe_Ratio <- (df$Portfolio_Return - df$RF)/df$Portfolio_SD

# let's verify that our dataframe is created correctly
head(df)
tail(df)
```

Now that the data has been created, let's create the charts in ggplot2
protip: download the ggplot2 cheat sheet https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf

```{r}
# Let's use GGPLOT to create our charts, it gives us more customization abilities
# for example we can encode the sharpe ratio as color along the curve

ggplot(df, aes(x=Portfolio_SD, y=Portfolio_Return)) +
  geom_point(aes(color=Sharpe_Ratio)) +
  xlab("Portfolio Standard Deviation") +
  ylab("Portfolio rate or return (in %)")
  
```

Let's create the second chart without the option of shorting
```{r}
# Careful, I'm using the filter function from the package dplyr, which is different from the standard
# one, therefore, load dyplr as library.

ggplot(filter(df, Large_Weight < 1, Small_Weight < 1), aes(x=Portfolio_SD, y=Portfolio_Return)) +
  geom_point(aes(color=Sharpe_Ratio)) +
  xlab("Portfolio Standard Deviation") +
  ylab("Portfolio rate or return (in %)")
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```